---
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.8in
fontsize: 11pt
title: "STA 445 S26 Assignment 4"
date: "`r Sys.Date()`"
output: pdf_document
---

Load all the packages you will use in this chunk.  Notice, I have set include = FALSE because I don't want to see you loading packages in the output.

```{r packages, include = FALSE}
library(tidyverse)
library(stringr)
```


## Problem 1

For the following regular expression, explain in words what it matches on. Then add test strings to demonstrate that it in fact does match on the pattern you claim it does.  Do at least 4 tests.  Make sure that your test set of strings has several examples that match as well as several that do not. Make sure to remove the `eval=FALSE` so that the code and output shows in your pdf.

a. This is detecting "a".

```{r}
        strings <- c("My dad is goofy", "The cat is large", "Dogs run quick",
                     "My silly brother")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, 'a') )
```
b. This is detecting "ab" next to each other in that order. 

```{r}
        
        strings <- c("Johnny is large", "I love laboratories", "What is today",
                     "Bang I am")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, 'ab') )
```

c. This is detecting for any instance of a or b

```{r}
        strings <- c("Johnny", "Who are you?", "Babadook", "Big silly gab",
                     "Bob is grand", "The chimney is big", "Hello")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '[ab]') )
```

d.  This is detecting for any instance of "ab" at the beginning of the string, where they are both lower case and in that order.

```{r}
        strings <- c("Dang it dude", "Who was born", "Banger", "Abdominals",
                     "Silly goose", "ab", "abstract", "rabid")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '^[ab]') )
```

e. This is detecting any digit plus the use of the letter a, capitalized or not. But they have to be spaced a part. That's why 12 A works but 12A does not. 

```{r}
        strings <- c("I like 28 architects", "The big deal", "1768", "12 a",
                     "12a", "12A", "12 A")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '\\d+\\s[aA]') )
```

f. This is detecting phrases that specifically have one or more numbers immediately followed by zero or more white spaces then a string beginning with a/A.

```{r}
        strings <- c("1284 big apples", "bongiorno to the 800", "what is the
                     Apple", "Aa 12", "28 a", "28 A", "29 Aa", "12 apples", 
                     "12 apples 12", "28a")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '\\d+\\s*[aA]') )
```

g. This is searching for zero or more repetitions of any character (always true).

```{r}
        strings <- c("Ok I am 22", "Bald or what", "1800 pigeons", "Singer.", 
                     "string string")
        data.frame( string = strings ) %>%
      mutate( result = str_detect(string, '.*') )
```

h. This is detecting for any phrase that, at the beginning of the string, has any two alphanumeric characters followed by the word bar. 

```{r}
        strings <- c("Gorgeous", "Bobby", "Singing", "Bigboy", "Lord 65 saint", 
                     "wabar")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '^\\w{2}bar') )
```

i. This is detecting for any character string that either has the string "foo.bar" in it or begins with any two alphanumeric characters followed by the word bar.  

```{r}
        strings <- c("foobar12bar", "foobar", "foo2bar", "foo.bar", "foo", 
                     "w8bar", "12 foo.bar")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '(foo\\.bar)|(^\\w{2}bar)') )
```
    
## Problem 2

The following file names were used in a camera trap study. The S number represents the site, P is the plot within a site, C is the camera number within the plot, the first string of numbers is the YearMonthDay and the second string of numbers is the HourMinuteSecond.

```{r}
    file.names <- c( 'S123.P2.C10_20120621_213422.jpg',
                     'S10.P1.C1_20120622_050148.jpg',
                 'S187.P2.C2_20120702_023501.jpg')
```
    
Produce a data frame with columns corresponding to the `site`, `plot`, `camera`, `year`, `month`, `day`, `hour`, `minute`, and `second` for these three file names. So we want to produce code that will create the data frame:
    
```{r, eval = FALSE}
     Site Plot Camera Year Month Day Hour Minute Second
     S123   P2    C10 2012    06  21   21     34     22
      S10   P1     C1 2012    06  22   05     01     48
      S187   P2     C2 2012    07  02   02     35     01
```

```{r}
data.frame(Camera = file.names) %>%
  separate(Camera, sep = "[.,_]", into = c("Site", "Plot", "Camera", "Date", 
                                           "Time", "File Type")) %>%
  mutate(Year = str_sub(Date, start = 1, end = 4)) %>%
  mutate(Month = str_sub(Date, start = 5, end = 6)) %>%
  mutate(Day = str_sub(Date, start = 7, end = 8)) %>%
  mutate(Hour = str_sub(Time, start = 1, end = 2)) %>%
  mutate(Minute = str_sub(Time, start= 3, end = 4)) %>%
  mutate(Second = str_sub(Time, start = 5, end = 6)) %>%
  select(Site, Plot, Camera, Year, Month, Day, Hour, Minute, Second)
```


3. The full text from Lincoln's Gettysburg Address is given below. Calculate the mean word length *Note: consider 'battle-field' as one word with 11 letters*).

```{r}
Gettysburg <- 'Four score and seven years ago our fathers brought forth on this continent, a new nation, conceived in Liberty, and dedicated to the proposition that all men are created equal. Now we are engaged in a great civil war, testing whether that nation, or any nation so conceived and so dedicated, can long endure. We are met on a great battle-field of that war. We have come to dedicate a portion of that field, as a final resting place for those who here gave their lives that that nation might live. It is altogether fitting and proper that we should do this. But, in a larger sense, we can not dedicate -- we can not consecrate -- we can not hallow -- this ground. The brave men, living and dead, who struggled here, have consecrated it, far above our poor power to add or detract. The world will little note, nor long remember what we say here, but it can never forget what they did here. It is for us the living, rather, to be dedicated here to the unfinished work which they who fought here have thus far so nobly advanced. It is rather for us to be here dedicated to the great task remaining before us -- that from these honored dead we take increased devotion to that cause for which they gave the last full measure of devotion -- that we here highly resolve that these dead shall not have died in vain -- that this nation, under God, shall have a new birth of freedom -- and that government of the people, by the people, for the people, shall not perish from the earth.'
```


```{r}
(Nopunctuation <- str_remove_all(Gettysburg, "[,,.,-]"))
```


```{r}
(Vectors <- str_split(Nopunctuation, pattern = "\\s+"))
(Vectors2 <- Vectors[[1]]) %>%
  str_length() %>%
  mean()
```

    